---
title: "Assignment 3 h"
author: "Harika"
date: "`r Sys.Date()`"
output: html_document
---
# QUESTION 
The file UniversalBank.csv contains data on 5000 customers of Universal Bank. The data include
customer demographic information (age, income, etc.), the customer’s relationship with the bank
(mortgage, securities account, etc.), and the customer response to the last personal loan campaign
(Personal Loan). Among these 5000 customers, only 480 (= 9.6%) accepted the personal loan that
was offered to them in the earlier campaign. In this exercise, we focus on two predictors: Online
(whether or not the customer is an active user of online banking services) and Credit Card
(abbreviated CC below) (does the customer hold a credit card issued by the bank), and the outcome
Personal Loan (abbreviated Loan below).
Partition the data into training (60%) and validation (40%) sets
```{r}
library(class)
library(caret)
library(dplyr)
library(e1071)

# Read the data
data <- read.csv("C:\\Users\\its me\\OneDrive - Kent State University\\Desktop\\Rhistory\\UniversalBank (1).csv")

# Display the first few rows of the data
head(data)

# Check the dimensions of the data
dim(data)
```


```{r}
# Partition the data into training (60%) and validation (40%) sets
set.seed(123)
trainIndex <- createDataPartition(data$Personal.Loan, p = 0.6, list = FALSE)
trainData <- data[trainIndex, ]
validationData <- data[-trainIndex, ]

# Display dimensions of training and validation sets
dim(trainData)
dim(validationData)
```

# QUESTION A:
Create a pivot table for the training data with Online as a column variable, CC as a row
variable, and Loan as a secondary row variable. The values inside the table should convey
the count. In R use functions melt() and cast(), or function table(). In Python, use panda
dataframe methods melt() and pivot().
```{r}
library(pander)

# Create a subset of relevant columns for analysis
subset <- trainData[c("CreditCard", "Personal.Loan", "Online")]

# Create a pivot table for the subset
pivot_table <- ftable(subset)

# Display the pivot table using pander
pandoc.table(pivot_table, style = "grid", split.tables = Inf)
```

# QUESTION B:
Consider the task of classifying a customer who owns a bank credit card and is actively using
online banking services. Looking at the pivot table, what is the probability that this customer
will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on
having a bank credit card (CC = 1) and being an active user of online banking services (Online
= 1)]
```{r}
# Check for conflicts
ls()

# If 'df' is not a conflict, reassign the dataframe
df <- data

# Convert 'Personal.Loan' to numeric if it's a factor
df$Personal.Loan <- as.numeric(as.character(df$Personal.Loan))

# Filter the data for customers with a bank credit card (CreditCard = 1) and active online banking (Online = 1)
subset_data <- subset(df, CreditCard == 1 & Online == 1)

# Calculate the probability of loan acceptance among the filtered customers
probability_loan_acceptance <- mean(subset_data$Personal.Loan, na.rm = TRUE)

# Print the result
cat("Probability of loan acceptance for customers with bank credit card and active online banking:", probability_loan_acceptance, "\n")


```

# QUESTION C:
Create two separate pivot tables for the training data. One will have Loan (rows) as a
function of Online (columns) and the other will have Loan (rows) as a function of CC.
```{r}
# Create subsets and pivot tables for other analyses (Online and CreditCard)
online_subset <- trainData[c("Personal.Loan", "Online")]
pivot_online <- ftable(online_subset)
pandoc.table(pivot_online, style = "grid", split.tables = Inf)

cc_subset <- trainData[c("Personal.Loan", "CreditCard")]
pivot_cc <- ftable(cc_subset)
pandoc.table(pivot_cc, style = "grid", split.tables = Inf)
```

# QUESTION D:
Compute the following quantities [P(A | B) means “the probability ofA given B”]:
i. P(CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors)
ii. P(Online = 1 | Loan = 1)
iii. P(Loan = 1) (the proportion of loan acceptors)
iv. P(CC = 1 | Loan = 0)
v. P(Online = 1 | Loan = 0)
vi. P(Loan = 0)
```{r}
# Calculate conditional probabilities
# i. P(CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors)
cc_loan1 <- pivot_cc[2, 2]
loan1 <- sum(pivot_cc[2, ])
CC_given_loan1 <- cc_loan1 / loan1
cat("Probability (CC = 1 | Loan = 1):", round(CC_given_loan1, 4), "\n")

# ii. P(Online = 1 | Loan = 1)
online_loan1 <- pivot_online[2, 2]
loan2 <- sum(pivot_online[2, ])
online_given_loan1 <- online_loan1 / loan2
cat("Probability (Online = 1 | Loan = 1):", round(online_given_loan1, 4), "\n")

# ii. P(Loan = 1) (the proportion of loan acceptors)
loan_acceptors <- sum(trainData$Personal.Loan == 1)
total <- nrow(trainData)
loan_acceptors1 <- loan_acceptors / total
cat("Probability (Loan = 1):", round(loan_acceptors1, 4), "\n")

# v. P(CC = 1 | Loan = 0)
cc_loan2 <- pivot_cc[1, 2]
loan3 <- sum(pivot_cc[1, ])
CC_notgiven_loan <- cc_loan2 / loan3
cat("Probability (CC = 1 | Loan = 0):", round(CC_notgiven_loan, 4), "\n")

# v. P(Online = 1 | Loan = 0)
online_loan2 <- pivot_online[1, 2]
loan4 <- sum(pivot_online[1, ])
online_notgiven_loan <- online_loan2 / loan4
cat("Probability (Online = 1 | Loan = 0):", round(online_notgiven_loan, 4), "\n")
 
#vi. P(Loan = 0)
nonloan_acceptors <- 1 - loan_acceptors1
cat("Probability (Loan = 0):", round(nonloan_acceptors, 4), "\n")
```

# QUESTION E:
Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC
= 1, Online = 1).
```{r}
# Calculate Naive Bayes probability
nb_prob <- (CC_given_loan1 * online_given_loan1 * loan_acceptors1) / 
           ((CC_given_loan1 * online_given_loan1 * loan_acceptors1) + 
            (CC_notgiven_loan * online_notgiven_loan * nonloan_acceptors))
cat("Probability (Personal.Loan  = 1 | CreditCard = 1, Online = 1):", round(nb_prob, 4), "\n")
```
# QUESTION F:
# Compare this value with the one obtained from the pivot table in (B). Which is a more
accurate estimate?

The provided values are probabilities related to loan acceptance for customers with a bank credit card (CC) and active online banking. 

In (B), the probability of loan acceptance for customers with a bank credit card and active online banking is given as 0.09297052.

In (E), the probability of loan acceptance given that the customer has a bank credit card (CC = 1) and active online banking (Online = 1) is given as 0.1106.

To compare these values, (E) appears to be a higher probability than the one obtained in (B). Therefore, based on the provided information, the estimate in (E) is higher and suggests a higher probability of loan acceptance for customers with a bank credit card and active online banking.



# QUESTION G:
Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)?
Run naive Bayes on the data. Examine the model output on training data, and find the entry
that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you
obtained in (E).

```{r}
# Build the Naive Bayes model
model <- naiveBayes(Personal.Loan ~ ., data = subset)
model
```


```{r}
# Calculate Naive Bayes probability using the model
result <- (0.327 * 0.643 * 0.092) / ((0.327 * 0.643 * 0.092) + (0.290 * 0.595 * 0.907))
cat("Prob (Loan = 1 | CC = 1, Online = 1):", round(result, 4))

```

The values provided for (E) and (G) are very close, with (E) being 0.1106 and (G) being 0.11.In a comparative sense, (E) and (G) can be considered nearly identical.