---
title: "FML Assignment 1"
author: "Harika"
date: "`r Sys.Date()`"
output: html_document
---


# QUESTION 1: 
# Download a dataset from the web. You may use any source, but specify the source in your code. Also ensure that the data has a mix of quantitative and qualitative (categorical) variables.

# QUESTION 2:
# Import the dataset into R

```{r}
# Load the dataset
house_prices <- read.csv("C:/Users/its me/OneDrive - Kent State University/Desktop/Rhistory/House_Prices.csv")
```

```{r}
str(house_prices)

```

# QUESTION 3:
# Print out descriptive statistics for a selection of quantitative and categorical variables.
```{r}
# Display the first few rows of the dataset
head(house_prices)

# Descriptive statistics for quantitative variables
summary(house_prices$SalePrice) 

# Descriptive statistics for categorical variables
table(house_prices$ FullBath)
```
#Descriptive statistics for quantitative variable "SalePrice":
The quantitative variable "SalePrice" exhibits the following descriptive statistics:

- Minimum (Min.): $34,900, denoting the lowest sale price in the dataset.

- First Quartile (1st Qu.): $130,000, indicating that 25% of the sale prices fall below this value.

- Median: $163,000, representing the middle value when all sale prices are arranged in ascending order.

- Mean: Approximately $183,108, serving as the average sale price.

- Third Quartile (3rd Qu.): $216,878, signifying that 75% of the sale prices are below this threshold.

- Maximum (Max.): $755,000, the highest sale price observed in the dataset.

#Descriptive statistics for categorical variable "FullBath":
Here are the descriptive statistics for the categorical variable "FullBath":

-Category 0: There are 5 instances where the number of full bathrooms is 0.

-Category 1: There are 405 instances where the number of full bathrooms is 1.

-Category 2: There are 467 instances where the number of full bathrooms is 2.

-Category 3: There are 23 instances where the number of full bathrooms is 3.


# QUESTION 4:
# Transform at least one variable. It doesn't matter what the transformation is.
```{r}
# Transform a variable 
house_prices$LogSalePrice <- log(house_prices$SalePrice)
```

# QUESTION 5:
# Plot at least one quantitative variable, and one scatterplot
```{r}
# Plotting a quantitative variable 
hist(house_prices$SalePrice, main = "Histogram of SalePrice", xlab = "SalePrice")

# Check for missing values in GarageArea and SalePrice
sum(is.na(house_prices$GarageArea))
sum(is.na(house_prices$SalePrice))

# Remove rows with missing values
house_prices <- na.omit(house_prices)

# Plotting a scatterplot 
plot(house_prices$GarageArea, house_prices$SalePrice, 
     main = "Scatterplot: GarageArea vs SalePrice",
     xlab = "GarageArea", ylab = "SalePrice")




```


